---
title: "Precept 4"
author: "Wei"
date: "February 25, 2016"
output: html_document
---

# Precept plan

0. Install `ggplot2` using the command `install.packages("ggplot2")`.
1. Finish base plotting lecture (and also mention `lines`)
3. Some general plotting options: colors, shapes, linetypes
2. Demonstrate plotting on the `diamonds` data set from `ggplot2`.


```{r data}
library(ggplot2)
data(diamonds)
head(diamonds)
```



### Barplots

Barplots take tables, or matrices of tables. In other words, they're useful for plotting categorical variables.

```{r barplots, fig.width=6, fig.height=4.5}
barplot(table(diamonds$color))
barplot(table(diamonds$color), space=0)
barplot(table(diamonds$color), names.arg=c("random", "labels", "for", "diamond", "color", "factor", "uhhh"))
barplot(table(diamonds$color), horiz=TRUE)
barplot(table(diamonds$color), col="firebrick")
barplot(table(diamonds$color), plot=FALSE)
```

### Histograms

Histograms are used to plot distributions.


```{r hist}
hist(diamonds$price)
hist(diamonds$price, breaks=250) #suggestion only
hist(diamonds$price, breaks=4)
hist(diamonds$price, main="Diamond prices", xlab="Price ($)", ylab="Count")
hist(diamonds$price, main="Diamond prices", xlab="Price ($)", ylab="Count", xlim=c(0,20000))
hist(diamonds$price, main="Diamond prices", xlab="Price ($)", ylab="Count", xlim=c(0,20000), col="rosybrown")

```

### Scatter plots

Scatter plots display two variables. A convention is to say that you are plotting thing-on-the-y-axis versus thing-on-the-x-axis.

```{r scatter plots}
plot(x=diamonds$carat, y=diamonds$price) #what do we see?
plot(x=diamonds$carat, y=diamonds$price, cex=3)
plot(x=diamonds$carat, y=diamonds$price, cex=0.2)
plot(x=diamonds$carat, y=diamonds$price, pch=5)
plot(x=diamonds$carat, y=diamonds$price, pch=5, cex=5)
plot(x=diamonds$carat, y=diamonds$price, type="l")
plot(x=seq(0, 10, length.out=500), y=seq(0, 10, length.out=500)^2, type="l")

plot(x=diamonds$carat, y=diamonds$price)
lines(seq(0,5,length.out=100), seq(0, 15000, length.out=100))

plot(x=diamonds$carat, y=diamonds$price)
abline(v=1, lty=3, col="hotpink", lwd=2)
abline(v=1.5, lty=3, col="hotpink")
abline(v=2, lty=3, col="hotpink")
abline(v=3, lty=3, col="hotpink")

class(diamonds$cut)
plot(x=diamonds$carat, y=diamonds$price, col=diamonds$color)
colors <- c("red", "darkorange", "lightgoldenrod", "seagreen", "cornflowerblue", "darkslateblue", "darkorchid")
plot(x=diamonds$carat, y=diamonds$price, col=colors[diamonds$color], pch=16, cex=0.25, xlab="Carat", ylab="Price ($)")
legend(x="bottomright", legend=levels(diamonds$color), col=colors, pch=16, title="Color")

```
